Bee visitation, pollination services, and plant yield in commodity and hybrid seed canola

### 2022-07-08 Manuscript/main.tex
Started on seed canola models
- Tried out a couple visitation models with/without Riley's data, with/without male bays included, and with/without male parameter included

- Conclusion: keep Riley's data but remove male term. Can mention that male bays aren't that different in the Methods section, but not include term in full model
Updated seed model claims

%Maps + climate graphs needed for publication
During 2015 and 2016, we also surveyed 35 hybrid seed canola fields (15 in 2015, 20 in 2016) near Lethbridge, Alberta, from June through August of each year, all of which had central-pivot irrigation
Seed fields were stocked with honey bee hives at a rate of 3.6 hives/ha, with apiaries stationed in the corners of fields, and were also stocked with LCBs in shelters throughout the field (1.3 or 2.6 shelters/ha).
% Shelters used either ``standard" or ``double"  LCB stocking ($\approx$50,000 vs. 100,000 cocoons/ha), and shelters were placed at either ``standard" or ``double" densities (1.3 vs. 2.6 shelters/ha) in each field.
% Double stocking rates were only used in fields with double shelter density, creating three combinations of LCB stocking: ``standard", ``double shelter", and ``double shelter and double stocking".
%Woudick 5,6 = 64 shelters (single), Woudick 2 = 113 shelters (double + bees), Woudick 3 = 128 shelters Graham 1 = 135 shelters

@@

In both fields, we established set of plots at varying distance from the sources of pollinators. 
In commodity fields, we established 1 m$^2$ plots at 5, 20, 100, and 500m, starting at the field margin closest to the set of honey bee hives %(271 total plots), 
and established plots starting at potential sources of natural pollinating insects in unstocked fields (forests, shrublands, or grasslands), but observed very few wild pollinators (Table \ref{tab:propVisitors}).
In seed canola fields, bays of hermaphroditic (herafter ``male") and female plants are typically planted in 1- and 6-m wide bays, respectively, and LCB shelters were located in the female bays (Figure \ref{fig:seedfieldPhoto}, \ref{fig:seedfieldPlots}).
We established plots at 5, 20, 100 (250 in 2016), and 400m into the field along a transect from the nearest set of honey bee hives, using pairs of plots at the edge of adjacent male and female bays at each distance (Figure \ref{fig:seedfieldPlots}).
We established another plot in the centre of the female bay at the 5m and 400m plot in each field to examine how visitation varied within the female bay.
We used a Nikon\texttrademark Laser 800S Rangefinder to measure the distance from each plot to the nearest shelter.
All surveys occurred on fair-weather days (median air temperature: 24.5$^\circ$C., range: 17--33) with no rain and as minimal wind as possible (less than 30km/hr).
Mean air temperature was similar between years in Grande Prairie ($\approx$ 22$^\circ$C), but Lethbridge was warmer in 2015 (23 vs. 28$^\circ$C).
Finally, we incorporated plot-level visitation data from Waytes (\citeyear{waytesMsc}; same years and locations) to more accurately estimate the effect of distance from shelter on visitation rates (647 total plots).
%Plots were marked with coloured stakes driven into the soil to locate them later in the season. %Full paper should have a table of climate variables. Windspeed needs sorting out (unit problems?)

Insect visitation and pollination data was collected once during the main canola bloom at each field (late June - late July), and plants were collected before harvest (mid - late August).
We recorded the number of insect visits that contacted the stigmas and anthers of flowers during 10 minutes of observation (5 min for seed fields during 2015), recorded the taxonomic group of the insect visitors, and counted the number of open, visitable flowers in each plot (petals had not yet started to dehisce, style had not elongated more than 3-4 millimeters beyond the anthers).
To assess pollen deposition, we collected stigmas from five random open flowers at each plot, mounted them in fuchsin gel \citep{beattie1971} on depression slides, and counted the pollen on each stigma using a Leica\texttrademark DME 13595 light microscope under 100x magnification (1294 commodity and 1050 seed canola stigmas).
At the end of the growing season, we collected three plants from the same plot and recorded the density of plants per m$^2$ (789 commodity canola plants, 582 seed canola plants).
After drying the plants, we weighed the entire plant, counted mature pods, and counted the number of flower pedicels to estimate total flower production. % at 50$^\circ$C
We randomly sampled five pods from each plant, and counted and weighed the seeds from each pod (3872 commodity canola pods, 2885 seed canola pods), then we threshed all the pods for each plant by hand, winnowed them using an air separator, and weighed the total mass of cleaned seeds.

\subsection{Analysis}
To examine how yield is influenced by visitation, we used piecewise structural equation models (pSEM, \citealp{shipley2009}).
Structural equation models are a set of linear models arranged in a causal network, which provides a framework for testing hypotheses about complex systems \citep{grace2012, lefcheck2015}.
Starting with a simple model (visitation $\rightarrow$ pollen deposition $\rightarrow$ fruit production $\leftarrow$ plant resources), we built a detailed model relating visitation to pollination to seed yield within a causal framework (Figures \ref{fig:commoditySEM} and \ref{fig:seedSEM}).
These models allowed us to model how visitation ultimately influences seed yield, while controlling for other confounding factors (e.g. distance into field, measurements taken at different fields).
Fruit set can suppress future vegetative growth and flower production \citep{stephenson1981}, but SEMs do not allow for causal loops \citep{grace2012}, so we included a path from fruit set to flower production as the closest approximation of this process.
The direction of the path could plausibly be reversed (implying that large flower production reduces the proportion of fruit set), but we considered only the first scenario, as \citet{sabbahi2006} showed that the converse process is more likely (where low fruit set causes increased flower production).

The underlying models of each pSEM were specified as generalized linear mixed-effects models (GLMMs, see Equation sets S1 and S2 for model specifications).
We used GLMMs to model processes that occur at different levels; for example, seed size-number trade-offs were modelled at the pod level, whereas pollen deposition was modeled at the flower level.
Random intercepts were used to model variance between fields, within-field (pollen, flower survival, seed count \& size), and within-plot (seed count \& size).
Continuous responses were modeled using log-normal (plant density, plant size), square-root normal (flower density), and exponential-normal distributions (weight per seed), while counts were modeled as a negative-binomial (honey bee visits, pollen grains per stigma, flowers per plant, seeds per pod) or zero-inflated negative-binomial distribution (LCB visits) with a log-link function.
Visitation models used $log$(time) as an ``offset" variable to account for differences in observation times.
Pod set (fraction of flowers that matured into pods) was modeled using a beta-binomial distribution with a logit-link function.
Interactions between predictors were evident in honey bee visitation rate (Year:Locality interaction) and plant density (Irrigation:Year) in commodity canola, and for LCB visitation (Tent Stocking:Distance from edge) in seed canola, and were included in each model (represented as Y-shaped path in Figures \ref{fig:commoditySEM} and \ref{fig:seedSEM}).

Structural equation models do not have paths between all variables, which may bias the model results if this independence assumption is incorrect; this requires testing a ``basis set" of independence claims using Shipley's d-separation criteria \citep{shipley2000, shipley2009}.
The initial commodity canola pSEM was misspecfied, as several missing paths were identified for the initial commodity model (Fisher's C = 167.5, df = 64, p$<<$0.001), but the updated pSEM that included missing paths was adequate (C = 81.6, df = 86, p = 0.62).
Similarly, the initial seed field pSEM also had several missing paths (C = 37.54, df = 56, p = 0.027), which we corrected (C = 58.10, df = 64, p = 0.68).
Basis sets were generated using the \texttt{dagitty} library in R 3.5.1.

@@

All component GLMMs of the pSEM were written in Stan 2.17.1 \citep{gelman2015}, and run using \texttt{rstan} 2.18.1.
We used informative normal priors ($\mu=0, \sigma=5$) for fixed effects, and gamma priors ($\alpha=1.05, \beta=1$) for the variance components.
We ran four separate chains with an adaptive phase of 1000 iterations, and a sampling phase of 1000 iterations, then checked for convergence of the chains ($\hat{R} \approx$1) and low autocorrelation within chains (high $N_{eff}$).
We assessed the adequacy of the underlying probability distribution functions of each model using posterior predictive checks \citep{gelman2013}, and found that the probability distributions were specified correctly (p$\sim$0.5) for all components except seed count.
This means that while the average seed counts (per plant) were unbiased, the variance within a plant was not. 
However, the negative binomial distribution had the closest posterior predictive checks of any other distribution. 
% p-values listed below (also in Tables S3 and S4) are \textit{posterior} p-values, representing the proportion of the posterior distribution that overlapped zero, with Z-scores (mean/standard deviation of posterior) showing the effect size.
% Predictions from models are median values of the posterior, representing the median prediction, while intervals displayed in figures are credible intervals (CIs), representing 95\% of the samples drawn from the posterior distribution (analogous to confidence intervals, see \citealt{gelman2013}).

@@

% %This could probably be reduced to a blurb
% Plant size and pollen deposition were the main drivers of yield in seed canola, while plant size was the main driver of yield in commodity canola.
% Distance from the field edge had a negative effect on honey bee visitation in both commodity and seed canola (Z=-3.7, -2.5, respectively), and unexpectedly, had a negative effect on leafcutter bee visitation (Z= -4.4), showing that overall visitation decreases with distance into the field.
% The path from leafcutter bee visitation to pollen deposition was strong in seed canola (Z=3.2), but there was no path from honey bee visitation to pollen deposition in either crop type.
% Pollen deposition increased fruit set (Z=3.1) and seeds per pod (Z=2.3) in seed canola, but was not present in commodity canola, showing that yield in seed canola, but not commodity canola, is strongly dependent on pollination.
% %The path from seeds per pod to weight per seed was negative in seed canola (Z=-15.0) but positive in commodity canola (Z=6.1) 
% Both crop types had strong negative paths from flower survival to flowers per plant (commodity Z=-12.5, seed Z=-12.1), indicating that high flower survival suppressed future flower production.
% Commodity canola had a similar weight per seed, regardless of the plant size, but in seed canola, plant size had a strong influence on seeds per pod (Z=5.6) and weight per seed (4.6).
% Finally, in both crop types, the paths leading from plant size were the dominant signal influencing fruit set (commodity Z=5.8; seed Z=19.9), but the path from pollen deposition to fruit set was only present in seed canola, showing that plant resources matter strongly for both crop types, but that increased pollen deposition only benefits seed canola.

The results of the underlying components of each pSEM are shown in Figures \ref{fig:commoditySEM} and \ref{fig:seedSEM}:

\begin{table}
\centering
\caption{Summary of main measured variables. Vegetative mass refers to the non-seed plant mass.}
\begin{tabular}{r|llllll|llllll}
                                       & \multicolumn{6}{l|}{Commodity Fields}           & \multicolumn{6}{l}{Seed Fields}                    \\
Parameter                              & Min  & Median & Mean   & Max   & SD     & N    & Min    & Median & Mean   & Max    & SD     & N     \\ \hline
Honey bee visitation (per min)      & 0    & 0      & 0.17   & 3.64  & 0.48   & 271  & 0      & 2      & 14.15  & 215    & 25.13  & 665   \\
Leafcutter bee visitation (per min) & -    & -      & -      & -     & -      & -    & 0      & 2      & 13.93  & 212    & 25.49  & 665   \\
Pollen (\# per stigma)                 & 0    & 155    & 293.32 & 3891  & 385.04 & 1355 & 0      & 7      & 21.76  & 578    & 42.53  & 1080  \\
Flowers per plant                      & 13   & 157    & 196.05 & 1094  & 150.87 & 789  & 26     & 362.50 & 461.33 & 2712   & 326.59 & 600   \\
Pods per plant                         & 5    & 112    & 143.17 & 892   & 114.57 & 789  & 10     & 244.50 & 299.41 & 1410   & 207.68 & 600   \\
Seeds per pod                          & 4.6  & 23.6   & 22.95  & 35.4  & 4.97   & 789  & 1.8    & 16.6   & 16.40  & 30.60  & 5.52   & 600   \\
Weight per 1000 seeds (g)              & 0.39 & 2.72   & 2.74   & 5.35  & 0.8    & 789  & 1.04   & 3.42   & 3.43   & 5.59   & 0.88   & 600  \\
Total vegetative mass (g)                    & 0.77 & 14.29  & 18.19  & 110.2 & 14.41  & 789  & 1.1849 & 25.20  & 30.32  & 144.32 & 21.16  & 600   \\
Total seed mass (g)                          & 0.01 & 5.46   & 6.87   & 47.9  & 5.97   & 789  & 0.02   & 7.54   & 9.50   & 60.77  & 7.90   & 600   
\end{tabular}
\end{table}

@@

Honey bee visitation decreased with distance away from the apiary, in both commodity and seed canola fields.
In commodity fields, honey bee visitation decreased with distance from the apiary at the field edge (Z=-3.7, p$=$0.0002, Figure \ref{fig:hbeeDist_commodity}), decreasing from 8.5 visits/hr at the edge of the field to 3.1 visits/hr at 100 m into the field (given a stocking level of 20 hives).
% Visitation rates were higher in both years in Lethbridge (2014: 9.3 visits/hr, 2015: 7.1), but were higher in 2015 than 2014 at Grande Prairie (Z=2.3, p$=$0.02), rising from 5 visits/hr at the field edge in 2014 to 17 visits/hr in 2015. % (at a stocking level of 20 hives). 
Greater stocking rates increased the honey bee visitation (Z=5.3, p$<$0.0001), with visitation at the edge of the field increasing from 14.5 visits/hr with an apiary of 20 hives to 22.7 visits/hr with an apiary of 40 hives, while unstocked field edges had an average of 1.8 visits/hr.
In seed canola fields, honey bee visitation also decreased with distance from the apiary (Z=-2.5, p$=$0.01), but the rate of decline was weaker (126 visits/hr at edge of field vs. 78 visits/hr at 100 m). 
Honey bee visitation was lower near leafcutter shelters (Z=3.9, p$=$0.001), dropping from 60 visits/hr at 10 m to 34 visits/hr at 2 m, likely due to competition with LCBs (Figure \ref{fig:hbeeDist_both}, \ref{fig:lbeeDist_both}).

LCB visitation declined rapidly with distance from their shelters (Z=-11.4, p$=$0.0001, 75\% reduction in 10 m), and unexpectedly, was higher at the edge of the field than the centre (Z=-4.4, p$=$0.0001, 114 visits/hr vs 28 visits/hr). %350 visit/hr at 2 m to 83 visit/hr at 10 m
However, there was no difference in visitation rates between the edge and the centre when leafcutter cocoon stocking was lower (Z=2.2, p$=$0.028), indicating that leafcutters at high stocking densities likely aggregate in shelters close to the edge of the field (Figure \ref{fig:hbeeDist_both}).
There was no difference in visitation rates between the male bay and the edge of the female bays (LCB: p$=$0.72, honey bee: p$=$0.32). 
However, LCB visitation was lower in the centre of the female bay (p$=$0.04, 51 vs. 29 visits/hr), and honey bee visitation was higher at the centre of the bay (p$=$0.008, 71 vs. 117 visits/hr, Figure \ref{fig:edgeCentVis_both}).
%Therefore, visitation by both species strongly decreases with distance from the edge of the field, decreases with distance from the leafcutter shelters in seed canola fields, and is done by different bees at the centre of the female bay.

@@

LCB visitation increased pollen deposition in seed canola fields, but honey bees did not increase pollen deposition in either commodity or seed canola fields.
Pollen deposition on stigmas was high in commodity canola (mean: 293 grains per stigma, SD: 385, range: 0--3981), but honey bee visitation did not effect deposition (p$=$0.77, Figure \ref{fig:hbeeVisPol}).
Pollen deposition slighly declined with distance from the edge of the field, but this relationship was weak (p$=$0.07) and only amounted to a 6\% average decrease (290 grains per stigma at the edge, 240 at field centre).
Flies were also common visitors in commodity fields (Table \ref{tab:propVisitors}), but had no effect on pollen deposition (p$=$0.577), so only honey bees and LCBs were considered as potential pollinators (other flower visitors were rare).
In seed canola, pollen deposition was much lower overall (mean: 22 grains per stigma, SD: 43, range: 0--578), LCB visitation had a strong positive effect (Z=3.2, p$=$0.002, Figure \ref{fig:visPol_both}, 12 grains/stigma at 6 visits/hr vs. 16 pollen grains at 60 visits/hr), and was lower at the centre of the field (Z=-4.4, p$=$0.0001, Figure \ref{fig:distCentPol}, 23 vs. 11 grains/stigma).
Pollen deposition was also lower at the centre of the female bay (Z=-4.5, p$=$0.0001, 23 vs. 13 grains/stigma) independent of visitation rate, suggesting that pollen may be lost from bees crossing from male bays, who likely visit flowers at the edge of the bay first \citep{pinnisch1990}. 
Finally, there was no effect of honey bees on pollen deposition in seed canola fields (Z=0.98, p$=$0.33), implying that most of the pollen deposition is the result of leafcutter visitation.

@@

Plant size had by far the strongest positive effect on the number of pods produced, and there was evidence of fruit set suppressing flower production.
Commodity canola plants produced fewer flowers (mean: 198, SD: 157, range: 13--1419) than seed canola plants (mean: 468, SD: 328, range: 26--2712), and also produced fewer pods per plant (mean: 143, SD: 114, range: 5--892) than seed canola plants (mean: 303, SD: 208, range: 10--1410).
In both commodity and seed canola, plant size had a strong positive effect on the total number of flowers per plant (commodity: Z=92.4, p$<$0.0001, seed: Z=82.7, p$<$0.0001, Figure \ref{fig:plSizeFlwCount}), as well as fruit set (commodity: Z=5.5, p$=$0.0002, seed: Z=19.9, p$<$0.0001, Figure \ref{fig:flwSurv_both}).
Fruit set had a strong negative influence on total flowers per plant (p$<$0.0001) in both commodity (Z=-12.5) and seed canola (Z=-12.1), indicating that high levels of fruit set suppressed future flower production.
Finally, commodity plants were also lower-yielding (mean: 6.8 g of seeds per plant, SD: 6.0, range: 0.01--47.9) than seed canola plants (mean: 9.6, SD: 7.9, range: 0.02--60.8).
However, the harvest index of commodity canola (mean: 0.26, SD: 0.08, range: 0.005--0.65) was higher than seed canola (mean: 0.23, SD: 0.09, range: 0.003--0.55), meaning that commodity plants produced more seeds per gram of biomass than seed canola plants, likely due to the lack of pollen limitation (see below).

>>>>

%Maps + climate graphs needed for publication
During 2015 and 2016, we also surveyed 35 hybrid seed canola fields (15 in 2015, 20 in 2016) near Lethbridge, Alberta, from June through August of each year, all of which had central-pivot irrigation.
In seed canola fields, bays of hermaphroditic (herafter ``male") and female plants are typically planted in 1- and 6-m wide bays, respectively. 
Seed fields were stocked with honey bee hives at a rate of 3.6 hives/ha, with apiaries stationed in the corners of fields, and were also stocked with LCB shelters at a rate of approximately 2 shelters/ha (Figure \ref{fig:seedfieldPhoto}, \ref{fig:seedfieldPlots}). % (1.3 or 2.6 shelters/ha).
% Shelters used either ``standard" or ``double"  LCB stocking ($\approx$50,000 vs. 100,000 cocoons/ha), and shelters were placed at either ``standard" or ``double" densities (1.3 vs. 2.6 shelters/ha) in each field.
% Double stocking rates were only used in fields with double shelter density, creating three combinations of LCB stocking: ``standard", ``double shelter", and ``double shelter and double stocking".
%Woudick 5,6 = 64 shelters (single), Woudick 2 = 113 shelters (double + bees), Woudick 3 = 128 shelters Graham 1 = 135 shelters

@@

In both field types, we established set of plots at varying distance from the sources of pollinators. 
In commodity fields, plots were located at 1 m$^2$ plots at 5, 20, 100, and 500 m (271 total plots), starting at the field edge closest to the set of honey bee hives or potential sources of natural pollinating insects in unstocked fields (forests, shrublands, or grasslands). 
We observed very few wild pollinators (Table \ref{tab:propVisitors}), so these were excluded from the analysis.
In seed fields, we established plots at 5, 20, 100 (250 m in 2016), and 400 m into the field along a transect from the nearest set of honey bee hives, using pairs of plots at the edge of adjacent male and female bays (Figure \ref{fig:seedfieldPlots}).
To examine within-bay variation in visitation, we established a plot at the centre of the female bay at the 5m and 400m plot in each field. 
Distances to nearest shelters were measured using a Nikon\texttrademark Laser 800S Rangefinder.
Finally, we incorporated plot-level visitation data from Waytes (\citeyear{waytesMsc}; same years and locations) to more accurately gauge the effect of distance from shelter on visitation rates (647 total plots).
All surveys occurred on fair-weather days (median air temperature: 24.5$^\circ$C., range: 17--33) with no rain and as minimal wind (less than 30km/hr).
% Mean air temperature was similar between years in Grande Prairie ($\approx$ 22$^\circ$C), but Lethbridge was warmer in 2015 (23 vs. 28$^\circ$C).
%Plots were marked with coloured stakes driven into the soil to locate them later in the season. %Full paper should have a table of climate variables. Windspeed needs sorting out (unit problems?)

Insect visitation and pollination data were collected once during the main canola bloom at each field (late June - late July), and plants were collected just prior to harvest (mid - late August).
We recorded the number of insect visits that contacted the stigmas and anthers of flowers during 10 minutes of observation (5 min for seed fields during 2015), recorded the identity of the visitor, and counted the number of open, visitable flowers in each plot (petals had not yet started to dehisce, style had not elongated more than 3-4 millimeters beyond the anthers).
To assess pollen deposition, we collected stigmas from five random flowers at each female plot, mounted them in fuchsin gel \citep{beattie1971} on depression slides, and counted the pollen on each stigma using a Leica\texttrademark DME 13595 light microscope under 100x magnification (1294 commodity and 1050 seed canola stigmas).
At the end of the growing season, we collected three plants from the same female plot and recorded the plant density per m$^2$ (789 commodity canola plants, 582 seed canola plants).
After drying the plants, we weighed the entire plant, counted mature pods to estimate pod set, and counted the number of flower pedicels to estimate total flower production. % at 50$^\circ$C
We estimated seed size and seeds per pod by averaging the seed count and weight from five pods on each plant.
Finally, we threshed all the pods for each plant by hand, winnowed them using an air separator, and weighed the total mass of cleaned seeds. % (3872 commodity canola pods, 2885 seed canola pods)

\subsection{Analysis}
To examine how yield is influenced by visitation, we used piecewise structural equation models (pSEM, \citealp{shipley2009}).
Structural equation models are a set of linear models arranged in a causal network, which provides a framework for testing hypotheses about complex systems \citep{grace2012, lefcheck2015}.
Starting with a simple model (visitation $\rightarrow$ pollen deposition $\rightarrow$ fruit production $\leftarrow$ plant resources), we built a detailed model relating visitation to pollination to seed yield within a causal framework (Figures \ref{fig:commoditySEM} and \ref{fig:seedSEM}).
These models let us model how visitation ultimately influences seed yield, while controlling for other confounding factors (e.g. distance into field, measurements taken at different fields).
Pod set can suppress future vegetative growth and flower production \citep{stephenson1981}, but SEMs do not allow for causal loops \citep{grace2012}, so we included a path from pod set to flower production as the closest approximation of this process.
The direction of the path could be reversed, implying that large flower production reduces the proportion of pod set, but we considered only the first scenario, as \citet{sabbahi2006} showed that low pod set increases flower production.

The underlying models of each pSEM were specified as generalized linear mixed-effects models (GLMMs, see Equation sets S1 and S2 for model specifications).
GLMMs can model processes that occur at different levels; for example, seed size was modelled at the plant level, while pollen deposition was modeled at the flower level.
% Random intercepts were used to model variance between fields, within-field (pollen, flower survival, seed count \& size), and within-plot (seed count \& size).
Random intercepts were used to model variance at the field level (and plot level, for data collected within plots); however, plot-level variance terms 
% Plot-level RE omitted from: commodity pollen, commodity flwSurv, commodity seedCount, seed plDens 
% Continuous responses were modeled using log-normal (plant density, plant size), square-root normal (flower density), and exponential-normal distributions (weight per seed), while counts were modeled as a negative-binomial (honey bee visits, pollen grains per stigma, flowers per plant, seeds per pod) or zero-inflated negative-binomial distribution (LCB visits) with a log-link function. (Move to appendix)
Visitation models used $log$(time) as an ``offset" variable to account for differences in observation times.
Plant and pollen samples were not collected from male bays, and there were no large differences in visitation rates for honey bees or LCBs (p = 0.32, 0.72, respectively) once other terms were included as covariates, so bay identity was excluded from further analysis.
% Pod set (fraction of flowers that matured into pods) was modeled using a beta-binomial distribution with a logit-link function.
% Interactions between predictors were evident in honey bee visitation rate (Year:Locality interaction) and plant density (Irrigation:Year) in commodity canola, and for LCB visitation (Tent Stocking:Distance from edge) in seed canola, and were included in each model (represented as Y-shaped path in Figures \ref{fig:commoditySEM} and \ref{fig:seedSEM}).

Structural equation models do not have paths between all variables, which may bias the model results if this independence assumption is incorrect; this requires testing a ``basis set" of independence claims using Shipley's d-separation criteria \citep{shipley2000, shipley2009}.
The initial commodity canola pSEM was misspecfied, as several missing paths were identified (Fisher's C = 167.5, df = 64, p$<<$0.001), but the updated pSEM was deemed adequate (C = 81.6, df = 86, p = 0.62).
Similarly, the initial seed field pSEM also had several missing paths (C = 105.6, df = 76, p = 0.01), which were added, making the updated pSEM adequate (C = 72.2, k = 72, p = 0.47).
Basis sets were generated using the \texttt{dagitty} library in R 3.5.1.

@@

All component GLMMs of the pSEM were written in Stan 2.17.1 \citep{gelman2015}, and run using \texttt{rstan} 2.18.1.
We used weakly informative normal priors ($\mu=0, \sigma=5$) for fixed effects and gamma priors ($\alpha=1, \beta=1$) for the variance components.
Four separate chains were run with an adaptive phase of 1000 iterations, and a sampling phase of 1000 iterations, then checked for convergence of the chains ($\hat{R} \approx$1) and low autocorrelation within chains (high $N_{eff}$).
We assessed the underlying probability distribution functions of each model using posterior predictive checks \citep{gelman2013}, and found that the probability distributions were adequately specified (p$\sim$0.5).
% This means that while the average seed counts (per plant) were unbiased, the variance within a plant was not. 
% However, the negative binomial distribution had the closest posterior predictive checks of any other distribution. 
% p-values listed below (also in Tables S3 and S4) are \textit{posterior} p-values, representing the proportion of the posterior distribution that overlapped zero, with Z-scores (mean/standard deviation of posterior) showing the effect size.
% Predictions from models are median values of the posterior, representing the median prediction, while intervals displayed in figures are credible intervals (CIs), representing 95\% of the samples drawn from the posterior distribution (analogous to confidence intervals, see \citealt{gelman2013}).

@@

% %This could probably be reduced to a blurb
% Plant size and pollen deposition were the main drivers of yield in seed canola, while plant size was the main driver of yield in commodity canola.
% Distance from the field edge had a negative effect on honey bee visitation in both commodity and seed canola (Z=-3.7, -2.5, respectively), and unexpectedly, had a negative effect on leafcutter bee visitation (Z= -4.4), showing that overall visitation decreases with distance into the field.
% The path from leafcutter bee visitation to pollen deposition was strong in seed canola (Z=3.2), but there was no path from honey bee visitation to pollen deposition in either crop type.
% Pollen deposition increased pod set (Z=3.1) and seeds per pod (Z=2.3) in seed canola, but was not present in commodity canola, showing that yield in seed canola, but not commodity canola, is strongly dependent on pollination.
% %The path from seeds per pod to weight per seed was negative in seed canola (Z=-15.0) but positive in commodity canola (Z=6.1) 
% Both crop types had strong negative paths from flower survival to flowers per plant (commodity Z=-12.5, seed Z=-12.1), indicating that high flower survival suppressed future flower production.
% Commodity canola had a similar weight per seed, regardless of the plant size, but in seed canola, plant size had a strong influence on seeds per pod (Z=5.6) and weight per seed (4.6).
% Finally, in both crop types, the paths leading from plant size were the dominant signal influencing pod set (commodity Z=5.8; seed Z=19.9), but the path from pollen deposition to pod set was only present in seed canola, showing that plant resources matter strongly for both crop types, but that increased pollen deposition only benefits seed canola.
% 
% The results of the underlying components of each pSEM are shown in Figures \ref{fig:commoditySEM} and \ref{fig:seedSEM}:
% 
% \begin{table}
% \centering
% \caption{Summary of main measured variables. Vegetative mass refers to the non-seed plant mass.}
% \begin{tabular}{r|llllll|llllll}
%                                        & \multicolumn{6}{l|}{Commodity Fields}           & \multicolumn{6}{l}{Seed Fields}                    \\
% Parameter                              & Min  & Median & Mean   & Max   & SD     & N    & Min    & Median & Mean   & Max    & SD     & N     \\ \hline
% Honey bee visitation (per min)      & 0    & 0      & 0.17   & 3.64  & 0.48   & 271  & 0      & 2      & 14.15  & 215    & 25.13  & 665   \\
% Leafcutter bee visitation (per min) & -    & -      & -      & -     & -      & -    & 0      & 2      & 13.93  & 212    & 25.49  & 665   \\
% Pollen (\# per stigma)                 & 0    & 155    & 293.32 & 3891  & 385.04 & 1355 & 0      & 7      & 21.76  & 578    & 42.53  & 1080  \\
% Flowers per plant                      & 13   & 157    & 196.05 & 1094  & 150.87 & 789  & 26     & 362.50 & 461.33 & 2712   & 326.59 & 600   \\
% Pods per plant                         & 5    & 112    & 143.17 & 892   & 114.57 & 789  & 10     & 244.50 & 299.41 & 1410   & 207.68 & 600   \\
% Seeds per pod                          & 4.6  & 23.6   & 22.95  & 35.4  & 4.97   & 789  & 1.8    & 16.6   & 16.40  & 30.60  & 5.52   & 600   \\
% Weight per 1000 seeds (g)              & 0.39 & 2.72   & 2.74   & 5.35  & 0.8    & 789  & 1.04   & 3.42   & 3.43   & 5.59   & 0.88   & 600  \\
% Total vegetative mass (g)                    & 0.77 & 14.29  & 18.19  & 110.2 & 14.41  & 789  & 1.1849 & 25.20  & 30.32  & 144.32 & 21.16  & 600   \\
% Total seed mass (g)                          & 0.01 & 5.46   & 6.87   & 47.9  & 5.97   & 789  & 0.02   & 7.54   & 9.50   & 60.77  & 7.90   & 600   
% \end{tabular}
% \end{table}

@@

Honey bee visitation decreased with distance away from the apiary, in both commodity and seed canola fields.
In commodity fields, honey bee visitation decreased from 8.5 visits/hr at the edge of the field to 3.1 visits/hr at 100 m into the field (p$=$0.0002, Figure \ref{fig:hbeeDist_commodity}, given a stocking rate of 20 hives).
% Visitation rates were higher in both years in Lethbridge (2014: 9.3 visits/hr, 2015: 7.1), but were higher in 2015 than 2014 at Grande Prairie (Z=2.3, p$=$0.02), rising from 5 visits/hr at the field edge in 2014 to 17 visits/hr in 2015. % (at a stocking level of 20 hives). 
Higher stocking rates increased honey bee visitation at the edge of the field from 14.5 visits/hr with an apiary of 20 hives to 22.7 visits/hr with an apiary of 40 hives (Z=5.3, p$<$0.0001), while unstocked field edges had an average of 1.8 visits/hr.
Honey bee visitation was much higher overall in seed canola fields, but still decreased from 126 visits/hr at edge of field to 78 visits/hr at 100 m from the apiary (Z=-2.5, p$=$0.01). 
Honey bee visitation was also lower near leafcutter shelters (Z=3.9, p$=$0.001), dropping from 60 visits/hr at 10 m to 34 visits/hr at 2 m, likely due to competition with LCBs (Figure \ref{fig:hbeeDist_both}, \ref{fig:lbeeDist_both}).

LCB visitation declined rapidly with distance from the nearest shelter, showing visitation rates of 350 visits/hr at 2 m declining to 83 visits/hr at 10 m (Z=-11.4, p$=$0.0001). 
Surprisingly, LCB visitation was also higher at the edge of the field than the centre (Z=-4.4, p$=$0.0001, 114 visits/hr vs 28 visits/hr).
% However, there was no difference in visitation rates between the edge and the centre when leafcutter cocoon stocking was lower (Z=2.2, p$=$0.028), indicating that leafcutters at high stocking densities likely aggregate in shelters close to the edge of the field (Figure \ref{fig:hbeeDist_both}).
% There was no difference in visitation rates between the male bay and the edge of the female bays (LCB: p$=$0.72, honey bee: p$=$0.32). 
LCB visitation was also lower in the centre of the female bays (p$=$0.04, 51 vs. 29 visits/hr), while honey bee visitation was higher (p$=$0.008, 71 vs. 117 visits/hr, Figure \ref{fig:edgeCentVis_both}), indicating that pollination is likely done by different species in different parts of the field.
%Therefore, visitation by both species strongly decreases with distance from the edge of the field, decreases with distance from the leafcutter shelters in seed canola fields, and is done by different bees at the centre of the female bay.

@@

LCB visitation increased pollen deposition in seed canola fields, but honey bees did not increase pollen deposition in either commodity or seed canola fields.
Pollen deposition on stigmas was high in commodity canola (mean: 293 grains per stigma, SD: 385, range: 0--3981), but honey bee visitation had no impact on this (p$=$0.77, Figure \ref{fig:hbeeVisPol}).
Pollen deposition slighly declined with distance from the edge of the field, but this relationship was weak (p$=$0.07) and only amounted to a 6\% average decrease (290 grains per stigma at the field edge vs. 240 at field centre).
% Flies were also common visitors in commodity fields (Table \ref{tab:propVisitors}), but had no effect on pollen deposition (p$=$0.577), so only honey bees and LCBs were considered as potential pollinators (other flower visitors were rare).
In seed canola, pollen deposition was much lower overall (mean: 22 grains per stigma, SD: 43, range: 0--578), and decreased from 23 to 11 grains/stigma from the edge to the centre of the field (Z=-4.4, p$=$0.0001, Figure \ref{fig:distCentPol}).
LCB visits increased pollen deposition from 12 grains/stigma at 6 visits/hr to 16 pollen grains at 60 visits/hr (Z=3.2, p$=$0.002, Figure \ref{fig:visPol_both}), but there was no effect of honey bee visitation (Z=0.98, p$=$0.33), implying that most of the pollen deposition in seed fields is the result of LCB visitation.
Pollen deposition also decreased from 23 grains/stigma at the edge of the female bay to 13 at the centre of the bay independent of visitation (Z=-4.5, p$=$0.0001), suggesting that pollen may be lost from bees crossing over from male bays \citep{pinnisch1990}.%, who likely visit flowers at the edge of the bay first.

@@

Plant size had by far the strongest positive effect on the number of pods produced, and there was evidence of pod set suppressing flower production.
Commodity canola plants produced fewer flowers (mean: 198, SD: 157, range: 13--1419) than seed canola plants (mean: 468, SD: 328, range: 26--2712), and also produced fewer pods per plant (mean: 143, SD: 114, range: 5--892 vs. mean: 303, SD: 208, range: 10--1410).
In both commodity and seed canola, plant size had a strong positive effect on the total number of flowers per plant (both p$<<$0.0001, Figure \ref{fig:plSizeFlwCount}), as well as pod set (both p$<$0.001, Figure \ref{fig:flwSurv_both}).
Pod set had a strong negative influence on total flowers per plant (p$<$0.0001) in both commodity and seed canola (both p$<<$0.0001), indicating that high pod set suppresses future flower production.
Finally, commodity plants were also lower-yielding (mean: 6.8 g of seeds per plant, SD: 6.0, range: 0.01--47.9) than seed canola plants (mean: 9.6, SD: 7.9, range: 0.02--60.8).
However, the harvest index of commodity canola (mean: 0.26, SD: 0.08, range: 0.005--0.65) was higher than seed canola (mean: 0.23, SD: 0.09, range: 0.003--0.55), meaning that commodity plants produced more seeds per gram of biomass than seed canola plants.%, likely due to the lack of pollen limitation.
